CTCF threshold analysis

### 2022-07-04 03_EDA_CTCFBSDB.Rmd
Rename

```{r settings}
# Project folder path
dir_data <- "/Users/mdozmorov/Documents/Work/GitHub/CTCF.dev"

# Input files
# Predicted data
fileNameIn1 <- file.path(dir_data, "data/allcomp.txt.gz")
# download.file("https://insulatordb.uthsc.edu/download/allcomp.txt.gz", fileNameIn1)
# Experimental data
fileNameIn2 <- file.path(dir_data, "data/CTCFBSDB_all_exp_sites_Sept12_2012.txt.gz")
# download.file("https://insulatordb.uthsc.edu/download/CTCFBSDB_all_exp_sites_Sept12_2012.txt.gz", fileNameIn2)
# Output files
fileNameOut1 <- file.path(dir_data, "results/hg18.CTCFBSDB.CTCF_predicted_human.RData")
fileNameOut1.1 <- file.path(dir_data, "results/hg18.CTCFBSDB.CTCF_predicted_human.bed")
fileNameOut2 <- file.path(dir_data, "results/mm8.CTCFBSDB.CTCF_predicted_mouse.RData")
fileNameOut2.1 <- file.path(dir_data, "results/mm8.CTCFBSDB.CTCF_predicted_mouse.bed")
# fileNameOut3 <- file.path(dir_data, "results/hg19.CTCFBSDB.CTCF_experimental_human.RData")
# fileNameOut3.1 <- file.path(dir_data, "results/hg19.CTCFBSDB.CTCF_experimental_human.bed")
# fileNameOut4 <- file.path(dir_data, "results/mm9.CTCFBSDB.CTCF_experimental_mouse.RData")
# fileNameOut4.1 <- file.path(dir_data, "results/mm9.CTCFBSDB.CTCF_experimental_mouse.bed")
```

@@

# Load EXPERIMENTAL data

```{r eval=FALSE}
# Read in data
mtx <- read_tsv(fileNameIn2)
```

```{r eval=FALSE}
# head(mtx, n = 100) %>% View()
colnames(mtx)
table(mtx$Species)
table(mtx$Build)
mtx$Build[mtx$Build == "galgal3"] <- "galGal3"
table(mtx$ValidationStatus)
table(mtx$CellType)
table(mtx$Occupancy)
summary(mtx$CTCF_Rank %>% as.numeric())
summary(mtx$Readcount %>% as.numeric())
table(mtx$ENCODE)
summary(mtx$ENCODE_Peak)
table(mtx$M1M2_class)
```

```{r eval=FALSE}
mtx_to_gr <- function(mtx = mtx, org = "Rhesus") {
  mtx_selected <- mtx[mtx$Species == org, ]
  gr <- GRanges(seqnames = mtx_selected$Chr, IRanges(start = mtx_selected$GenomeBeginLoc, end = mtx_selected$GenomeEndLoc))

@@

organisms <- unique(mtx$Species)
# organisms <- organisms[organisms != c("Rhesus")]
# for(i in organisms) {
#   print(i)
#   gr <- mtx_to_gr(mtx, org = i)
#   # Output files
#   fileNameOut1 <- file.path(dir_data, paste0("results/CTCF_", i, "_CTCFBSDB_experimental.RData"))
#   # Save as RData object. 
#   # save(list = mtx_to_gr(mtx = mtx_human, genome_id = "hg19"), file = fileNameOut1)
# }

>>>>

```{r settings}
# Project folder path
dir_project <- "/Users/mdozmorov/Documents/Work/GitHub/CTCF.dev"

# Input files
# Predicted data
fileNameIn1 <- file.path(dir_project, "data/allcomp.txt.gz")
if (!file.exists(fileNameIn1)) {
  download.file("https://insulatordb.uthsc.edu/download/allcomp.txt.gz", fileNameIn1)
}
# Experimental data
fileNameIn2 <- file.path(dir_project, "data/CTCFBSDB_all_exp_sites_Sept12_2012.txt.gz")
# download.file("https://insulatordb.uthsc.edu/download/CTCFBSDB_all_exp_sites_Sept12_2012.txt.gz", fileNameIn2)
# Output files
fileNameOut1 <- file.path(dir_project, "RData", "hg18.CTCFBSDB.CTCF_predicted_human.RData")
fileNameOut1.1 <- file.path(dir_project, "RData", "hg18.CTCFBSDB.CTCF_predicted_human.bed")
fileNameOut2 <- file.path(dir_project, "RData", "mm8.CTCFBSDB.CTCF_predicted_mouse.RData")
fileNameOut2.1 <- file.path(dir_project, "RData", "mm8.CTCFBSDB.CTCF_predicted_mouse.bed")
# fileNameOut3 <- file.path(dir_project, "results/hg19.CTCFBSDB.CTCF_experimental_human.RData")
# fileNameOut3.1 <- file.path(dir_project, "results/hg19.CTCFBSDB.CTCF_experimental_human.bed")
# fileNameOut4 <- file.path(dir_project, "results/mm9.CTCFBSDB.CTCF_experimental_mouse.RData")
# fileNameOut4.1 <- file.path(dir_project, "results/mm9.CTCFBSDB.CTCF_experimental_mouse.bed")
```

@@

```{r eval=FALSE}
# Load EXPERIMENTAL data

# Read in data
mtx <- read_tsv(fileNameIn2)
```

```{r eval=FALSE}
## EDA

# head(mtx, n = 100) %>% View()
colnames(mtx)
table(mtx$Species)
table(mtx$Build)
mtx$Build[mtx$Build == "galgal3"] <- "galGal3"
table(mtx$ValidationStatus)
table(mtx$CellType)
table(mtx$Occupancy)
summary(mtx$CTCF_Rank %>% as.numeric())
summary(mtx$Readcount %>% as.numeric())
table(mtx$ENCODE)
summary(mtx$ENCODE_Peak)
table(mtx$M1M2_class)
```

```{r eval=FALSE}
## GRanges

mtx_to_gr <- function(mtx = mtx, org = "Rhesus") {
  mtx_selected <- mtx[mtx$Species == org, ]
  gr <- GRanges(seqnames = mtx_selected$Chr, IRanges(start = mtx_selected$GenomeBeginLoc, end = mtx_selected$GenomeEndLoc))

@@

organisms <- unique(mtx$Species)
# organisms <- organisms[organisms != c("Rhesus")]
# for(i in organisms) {
#   print(i)
#   gr <- mtx_to_gr(mtx, org = i)
#   # Output files
#   fileNameOut1 <- file.path(dir_project, paste0("results/CTCF_", i, "_CTCFBSDB_experimental.RData"))
#   # Save as RData object. 
#   # save(list = mtx_to_gr(mtx = mtx_human, genome_id = "hg19"), file = fileNameOut1)
# }

### 2022-07-04 05_EDA_liftOver_mm.Rmd
Rename

```{r settings}
# Project folder path
dir_data <- "/Users/mdozmorov/Documents/Data/GoogleDrive/CTCF.dev/CTCF_liftover/"
dir_project <- "/Users/mdozmorov/Documents/Work/GitHub/CTCF.dev"
# Input files
fileNameIn1 <- file.path(dir_data, "sample_annotation.csv")
# Output files
fileNameOut1 <- file.path(dir_project, "results", "Figure_liftOverJaccard.svg")
```

@@

```{r}
# Adjust clustering method
# if(genome_assembly == "hg38") {
#   clust_method <- "euclidean"
# }
# if(genome_assembly == "mm10") {
#   clust_method <- "correlation"
# }
clust_method <- "euclidean"
# Save the plot
col3 <- colorRampPalette(c('white', 'white','white','white','white','white','white','orangered'))(20)
# all.equal(colnames(mtx_to_plot), sample_annotation$Name)
mydf <- data.frame(Assembly = sample_annotation$Assembly)
rownames(mydf) <- sample_annotation$Name
# png("man/figures/excluderanges_hg38_jaccard.png", width = 1000, height = 900, res = 200)
p <- pheatmap(data.matrix(mtx_to_plot), cluster_cols = T, cluster_rows = T, color = col3,
         clustering_method = "ward.D",# "ward.D", 
         clustering_distance_rows = clust_method, 
         clustering_distance_cols = clust_method, 
         annotation_row = mydf, annotation_colors = list(Assembly = c(hg18 = mycols[2], hg19 = mycols[4], hg38 = mycols[5], T2T = mycols[1])),
         treeheight_row = 40,
         treeheight_col = 0, 
         display_numbers = TRUE)

>>>>

```{r settings}
# Project folder path
dir_data <- "/Users/mdozmorov/Documents/Data/GoogleDrive/CTCF.dev/CTCF_liftover/"
dir_project <- "/Users/mdozmorov/Documents/Work/GitHub/CTCF.dev"
# Input files
fileNameIn1 <- file.path(dir_data, "sample_annotation_mm.csv")
# Output files
fileNameOut1 <- file.path(dir_project, "results", "Figure_liftOverJaccard_mm.svg")
```

@@

```{r}
# Adjust clustering method
# if(genome_assembly == "hg38") {
#   clust_method <- "euclidean"
# }
# if(genome_assembly == "mm10") {
#   clust_method <- "correlation"
# }
clust_method <- "euclidean"
# Save the plot
col3 <- colorRampPalette(c('white', 'white','white','white','white','white','white','orangered'))(20)
# all.equal(colnames(mtx_to_plot), sample_annotation$Name)
mydf <- data.frame(Assembly = sample_annotation$Assembly)
rownames(mydf) <- sample_annotation$Name
# png("man/figures/excluderanges_hg38_jaccard.png", width = 1000, height = 900, res = 200)
p <- pheatmap(data.matrix(mtx_to_plot), cluster_cols = T, cluster_rows = T, color = col3,
         clustering_method = "ward.D",# "ward.D", 
         clustering_distance_rows = clust_method, 
         clustering_distance_cols = clust_method, 
         annotation_row = mydf, annotation_colors = list(Assembly = c(mm9 = mycols[2], mm10 = mycols[4], mm39 = mycols[5])),
         treeheight_row = 40,
         treeheight_col = 0, 
         display_numbers = TRUE)

### 2022-07-04 06_FIMO_EDA.Rmd
Rename

```{r settings}
dir_data <- "/Users/mdozmorov/Documents/Data/GoogleDrive/CTCF.dev/merlot"
# Motifs
motif_human  <- c("CIS_BP_2.00_Homo_sapiens", "CTCFBSDB_PWM", "HOCOMOCOv11_core_HUMAN_mono_meme_format", "JASPAR2022_CORE_vertebrates_non_redundant_v2", "jolma2013", "SwissRegulon_human_and_mouse")
motif_mouse  <- c("CIS_BP_2.00_Mus_musculus", "CTCFBSDB_PWM", "HOCOMOCOv11_core_MOUSE_mono_meme_format", "JASPAR2022_CORE_vertebrates_non_redundant_v2", "jolma2013", "SwissRegulon_human_and_mouse")
# Genomes
genome_human <- c("T2T", "hg38", "hg19")
genome_mouse <- c("mm10", "mm9")
```

```{r hg18, eval=FALSE}
# Human data hg18
pvals <- rbind()
pval_threshold <- 1e-6
for (gen in c("hg18", genome_human)) {
  for (mot in "JASPAR2022_MA0139.1") {
    print(paste(gen, mot, sep = "."))
    mtx <- read_tsv(fileNameIn, col_names = FALSE, col_types = c("ciicdciddc"))
    fileNameIn <- file.path(dir_data, paste(gen, mot, "bed", sep = "."))
    print(paste("Number of regions more significant than the p-value threshold:", round(sum(mtx$X8 < pval_threshold), digits = 5) ))
    print(paste("Proportion of regions more significant than the p-value threshold:", round(sum(mtx$X8 < pval_threshold) / nrow(mtx), digits = 5) ))
    pvals <- rbind(pvals, cbind(-log10(mtx$X8), rep(gen, nrow(mtx))))
    # pvals$pvals <- c(pvals$pvals, -log10(mtx$X8))
    # pvals$assembly <- c(pvals$assembly, rep(gen, nrow(mtx)))
  }
}
pvals <- as.data.frame(pvals)
pvals[, 1] <- as.numeric(pvals[, 1])
colnames(pvals) <- c("-log10(p-value)", "Assembly")

@@

```{r hgall}
pvals <- rbind()
pval_threshold <- 1e-6
for (gen in "hg38") {
  for (mot in motif_human) {
    print(paste(gen, mot, sep = "."))
    fileNameIn <- file.path(dir_data, paste(gen, mot, "bed", sep = "."))
    mtx <- read_tsv(fileNameIn, col_names = FALSE, col_types = c("ciicdciddc"))
    print(paste("Number of regions more significant than the p-value threshold:", round(sum(mtx$X8 < pval_threshold), digits = 5) ))
    print(paste("Proportion of regions more significant than the p-value threshold:", round(sum(mtx$X8 < pval_threshold) / nrow(mtx), digits = 5) ))
    pvals <- rbind(pvals, cbind(-log10(mtx$X8), rep(mot, nrow(mtx))))
  }
}
pvals <- as.data.frame(pvals)
pvals[, 1] <- as.numeric(pvals[, 1])
colnames(pvals) <- c("-log10(p-value)", "MotifDB")

@@

fileNameOut1 <- file.path(dir_data, "Figure_human_pvalues.svg")
ggsave(filename = fileNameOut1, width = 6.4, height = 1.6)

@@

```{r mmall}
pvals <- rbind()
pval_threshold <- 1e-6
for (gen in "mm10") {
  for (mot in motif_mouse) {
    print(paste(gen, mot, sep = "."))
    fileNameIn <- file.path(dir_data, paste(gen, mot, "bed", sep = "."))
    mtx <- read_tsv(fileNameIn, col_names = FALSE, col_types = c("ciicdciddc"))
    print(paste("Number of regions more significant than the p-value threshold:", round(sum(mtx$X8 < pval_threshold), digits = 5) ))
    print(paste("Proportion of regions more significant than the p-value threshold:", round(sum(mtx$X8 < pval_threshold) / nrow(mtx), digits = 5) ))
    pvals <- rbind(pvals, cbind(-log10(mtx$X8), rep(mot, nrow(mtx))))
  }
}
pvals <- as.data.frame(pvals)
pvals[, 1] <- as.numeric(pvals[, 1])
colnames(pvals) <- c("-log10(p-value)", "MotifDB")

@@

fileNameOut1 <- file.path(dir_data, "Figure_mouse_pvalues.svg")
# ggsave(filename = fileNameOut1, width = 7, height = 3)
dev.off()
svg(filename = fileNameOut1, width = 7, height = 3)
print(p)
dev.off()
```

>>>>

```{r settings}
dir_data <- "/Users/mdozmorov/Documents/Data/GoogleDrive/CTCF.dev/merlot"
dir_project <- "/Users/mdozmorov/Documents/Work/GitHub/CTCF.dev"
# Motifs
motif_human  <- c("CIS_BP_2.00_Homo_sapiens", "CTCFBSDB_PWM", "HOCOMOCOv11_core_HUMAN_mono_meme_format", "JASPAR2022_CORE_vertebrates_non_redundant_v2", "Jolma2013", "SwissRegulon_human_and_mouse")
motif_mouse  <- c("CIS_BP_2.00_Mus_musculus", "CTCFBSDB_PWM", "HOCOMOCOv11_core_MOUSE_mono_meme_format", "JASPAR2022_CORE_vertebrates_non_redundant_v2", "Jolma2013", "SwissRegulon_human_and_mouse")
# Genomes
genome_human <- c("T2T", "hg38", "hg19")
genome_mouse <- c("mm10", "mm9", "mm39")
```

```{r hg18, eval=FALSE}
# Human data hg18
pvals <- rbind()
pval_threshold <- 1e-6
for (gen in c("hg18", genome_human)) {
  for (mot in "JASPAR2022_MA0139.1") {
    print(paste(gen, mot, sep = "."))
    fileNameIn <- file.path(dir_project, paste(gen, mot, "bed", sep = "."))
    mtx <- read_tsv(fileNameIn, col_names = FALSE, col_types = c("ciicdciddc"))
    print(paste("Number of regions more significant than the p-value threshold:", round(sum(mtx$X8 < pval_threshold), digits = 5) ))
    print(paste("Proportion of regions more significant than the p-value threshold:", round(sum(mtx$X8 < pval_threshold) / nrow(mtx), digits = 5) ))
    pvals <- rbind(pvals, cbind(-log10(mtx$X8), rep(gen, nrow(mtx))))
    # pvals$pvals <- c(pvals$pvals, -log10(mtx$X8))
    # pvals$assembly <- c(pvals$assembly, rep(gen, nrow(mtx)))
  }
}
pvals <- as.data.frame(pvals)
pvals[, 1] <- as.numeric(pvals[, 1])
colnames(pvals) <- c("-log10(p-value)", "Assembly")

@@

```{r hgall}
pvals <- rbind()
pval_threshold <- 1e-6
for (gen in "hg38") {
  for (mot in motif_human) {
    print(paste(gen, mot, sep = "."))
    fileNameIn <- file.path(dir_project, "RData", paste(gen, mot, "bed", sep = "."))
    mtx <- read_tsv(fileNameIn, col_names = FALSE, col_types = c("ciicdciddc"))
    print(paste("Number of regions more significant than the p-value threshold:", round(sum(mtx$X8 < pval_threshold), digits = 5) ))
    print(paste("Proportion of regions more significant than the p-value threshold:", round(sum(mtx$X8 < pval_threshold) / nrow(mtx), digits = 5) ))
    pvals <- rbind(pvals, cbind(-log10(mtx$X8), rep(mot, nrow(mtx))))
  }
}
pvals <- as.data.frame(pvals)
pvals[, 1] <- as.numeric(pvals[, 1])
colnames(pvals) <- c("-log10(p-value)", "MotifDB")

@@

fileNameOut1 <- file.path(dir_project, "results", "Figure_human_pvalues.svg")
ggsave(filename = fileNameOut1, width = 6.4, height = 1.6)

@@

```{r mmall}
pvals <- rbind()
pval_threshold <- 1e-6
for (gen in "mm10") {
  for (mot in motif_mouse) {
    print(paste(gen, mot, sep = "."))
    fileNameIn <- file.path(dir_project, "RData", paste(gen, mot, "bed", sep = "."))
    mtx <- read_tsv(fileNameIn, col_names = FALSE, col_types = c("ciicdciddc"))
    print(paste("Number of regions more significant than the p-value threshold:", round(sum(mtx$X8 < pval_threshold), digits = 5) ))
    print(paste("Proportion of regions more significant than the p-value threshold:", round(sum(mtx$X8 < pval_threshold) / nrow(mtx), digits = 5) ))
    pvals <- rbind(pvals, cbind(-log10(mtx$X8), rep(mot, nrow(mtx))))
  }
}
pvals <- as.data.frame(pvals)
pvals[, 1] <- as.numeric(pvals[, 1])
colnames(pvals) <- c("-log10(p-value)", "MotifDB")

@@

fileNameOut1 <- file.path(dir_project, "results", "Figure_mouse_pvalues.svg")
# ggsave(filename = fileNameOut1, width = 7, height = 3)
dev.off()
svg(filename = fileNameOut1, width = 6.4, height = 1.6)
print(p)
dev.off()
```

### 2022-07-04 07_CTCF_Threshold.Rmd
Rename

# Input files
fileNameIn1 <- file.path(dir_project, "RData", "mm10.MA0139.1.bed")
fileNameIn2 <- file.path(dir_project, "data", "mm10-CTCF.bed")
# Output files
fileNameOut1 <- file.path(dir_project, "results", "Figure_mouse_pvalues_threshold.svg")
```

>>>>

# Input files
fileNameIn1 <- file.path(dir_project, "RData", "hg38.MA0139.1.bed")
fileNameIn2 <- file.path(dir_project, "data", "GRCh38-CTCF.bed")
# Output files
fileNameOut1 <- file.path(dir_project, "results", "Figure_human_pvalues_threshold.svg")
```
